---
title: "TBhotspotsPostratification"

subtitle: "Post stratification of TB prevalence"

author: |
  | McEwen Khundi
  |
  | London School of Hygiene & Tropical Medicine, London, UK
  | Malawi-Liverpool-Wellcome Clinical Research Programme, Blantyre, Malawi
  |

date: | 
  | `r format(Sys.time(), "%B %d, %Y")`
  |
  | Table of Contents:
output: 
  html_document:
    df_print: paged
    theme: spacelab
    toc: true
---

<br>

## 1. Backgound


<br>

## 2. Set-up

Load all required packages for analysis.

```{r setup, message=FALSE}
library(tidyverse)    #for data manipulation
library(pmthemes)     #for ggplot themes
library(knitr)        #for tables
library(brms)
library(here)
library(gt)
library(janitor)
library(brms)
library(tidybayes)
```

<br>

## 3. Import datasets

Import data required for the analysis.

```{r import}
ind_xray_hts_subdata_tbcse <- readRDS(here::here("data/ind_xray_hts_subdata_tbcse.rds"))

wp_2019 <- readRDS(here::here("data/wp_2019.rds")) %>% ungroup() %>% mutate(p_pop=N_wp/sum(N_wp)) %>% mutate(cluster=as.factor(paste0("c",clust)))


```

<br>

## 4. Tidying

```{r}
#Checking data with missing values
ind_xray_hts_subdata_tbcse %>%
  filter(rowSums(across(everything(), ~is.na(.x))) > 0 )

age_levels2 <- c("16-19","20-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65-69","70+") # includes up to 70
age_levels_hack <- c(18,22,27,32,37,42,47,52,57,62,67,72)
```

<br>

## 5. Model

```{r tidy}
m1 <- readRDS(here("data/m1.rds"))
m1
plot(m1)
# Prior names
# get_prior(formula =tbcase ~ sex + age + (1|cluster), 
#           data = ind_xray_hts_subdata_tbcse , family = bernoulli())
# 
# # Priors
# p1 <- c(
#   prior(normal(0,10), class=Intercept), # so this is my estimate for intercept at base case??? 
#   prior(normal(1,10), class="b"), #log-odds for women compared to men (see graph at top, log-odds 1 roughly equvilent to doubling)
#   prior(cauchy(0,1), sd) # can't remember/not sure what this is
# )
# 
# m1 <- brm(tbcase ~ sex + age + (1|cluster), 
#           data=ind_xray_hts_subdata_tbcse, 
#           family=bernoulli(), 
#           prior=p1,
#           chains=4,
#           cores=4)
# 
# saveRDS(m1,here::here("data/m1.rds"))
```
## 6. Scaffold df
Create df for predictions

Create a scaffold df - one row per every combo of age, sex, cluster (one age at midpoint of every age group for now)

```{r}
# Create a scaffold df - one row per every combo of age, sex, cluster (one age at midpoint of every age group for now)
newdata_blank <- list(cluster=paste0("c",c(1:72)), age=c(18,22,27,32,37,42,47,52,57,62,67,72), sex=c("Male", "Female")) %>% expand.grid() %>% as_tibble()


a <- readRDS(here("data/a.rds"))
a <- tidybayes::add_fitted_draws(newdata=newdata_blank, model=m1) 

#saveRDS(a, file = here::here("data/a.rds"))
```
So this makes sense for a summary of probability of being TB positive at any age / sex / cluster category

```{r}
a %>% group_by(age,sex,cluster) %>% median_qi(.value)
```
And we can use this to draw the graph we like (although NB. this graph is less good as age is one of 9 levels, not over every level)
```{r}
a %>% group_by(age,sex,cluster) %>% median_qi(.value) %>%
  ggplot() +
  geom_path(aes(x=age, y=.value*100000, group=cluster, colour=sex), alpha=0.2) +
  facet_grid(sex~.) +
  scale_color_brewer(palette = "Set1") +
  #scale_colour_manual(values = c("cyan4", "darkorange")) +
  labs(y="TB prevalence per 100000",
       x="Age (years)") +
  theme_bw() +
  theme(legend.position = "none")
```
BUT there are two problems with "overall" TB prevalence here. (a) the CI is too wide, (b) I am pretty sure this is overall HIV prevalence (?probability) for a population where there is one person at every age-sex-cluster level.

See below.  TB prevalence 179.22 per 100,000 (too high, it's 13% in reality) and CI goes from 33.64 per 100,000 to 672.91 per 100,000.  

To me this CI makes sense if you had to guess probability of any one person picked at random in Blantyre being TB positive, but it doesn't make sense as a population prevalence (because not everyone in population is going to have extreme values).

```{r}
a %>% ungroup() %>% median_qi(.value)

0.001792206*100000;	0.0003364706*100000;	0.006729115*100000
```
So, not claiming this is sensible, but it is possible to take the fitted_draws df and gather a summary of each "hypothetical" one-person-per-age-sex-category population. Then can median_qi the summary...


```{r}
a %>% ungroup() %>% mutate(set=rep(1:4000, times=nrow(newdata_blank))) %>% # the defaults in add_fitted_draws seem to have 4000 draws per age-sex-clust category
  group_by(set) %>% 
  summarise(prev=median(.value)) %>% # can take median here as there is one "person" (row) per age-sex-cluster category, so nothing to weight
  median_qi(prev)

0.001782939*100000;	0.001016186*100000;	0.002691657*100000
```
^^ Still 178.29 (which I think is wrong), but CI is much smaller ( 101.6186 to 269.1657)

## 7. A scaffold from the data itself

```{r, message=F}
newdata_scale <- ind_xray_hts_subdata_tbcse %>% 
  mutate(age = cut(age, breaks=c(15,20,25,30,35,40,45,50,55,60,65,70,Inf), right=F,  labels=age_levels_hack)) %>%
  mutate(age=as.numeric(as.character(age))) %>%
  group_by(age,sex,cluster) %>% summarise(n=n()) %>% 
  right_join(newdata_blank) %>%
  mutate(n=case_when(is.na(n)~0L, T ~n)) # to make NAs into zeros
newdata_scale # so this is a dataframe with the number of people in each age-sex-cluster group ("n")

rm(a)

b <- readRDS(here("data/b.rds"))
# b <- tidybayes::add_fitted_draws(newdata=newdata_scale, model=m1) 
# saveRDS(b, file = here::here("data/b.rds"))
b %>% ungroup() %>% median_qi(.value) # same as above so far - just one person for each row (have ignored "n")

0.001792206*100000;	0.0003364706*100000;	0.006729115*100000
```
Now combine with how many people are in each age-sex-cluster category (there are 4000 draws per level, so 4000 different hypothetical populations)

```{r}

b %>% ungroup() %>% 
  mutate(set=rep(1:4000, times=nrow(newdata_blank))) %>%
  group_by(set) %>%
  mutate(p_pop=n/sum(n)) %>% ###Every sum(n) is the same try to find out from Rachael
  mutate(pp =.value * p_pop) %>%
  summarise(prev=sum(pp)) # Sothis is the overall prevalence, in 4000 different iterations of draws

b %>% ungroup() %>% 
  mutate(set=rep(1:4000, times=nrow(newdata_blank))) %>%
  group_by(set) %>%
  mutate(p_pop=n/sum(n)) %>%
  mutate(pp =.value * p_pop) %>%
  summarise(prev=sum(pp)) %>% # Sothis is the overall prevalence, in 4000 different iterations of draws
  median_qi(prev)

0.002192346*100000;	0.001494808*100000;	0.003076053*100000;
```
So this estimate ^ (219) is very similar to data, with a v small CI (149 -  307). 

Could also do something similar to calculate cluster-specific prevalence; given the age and sex distribution in each cluster (in the tested dataset)

```{r, message=F, warning=F}
c <- b %>% ungroup() %>% 
  mutate(set=rep(1:4000, times=nrow(newdata_blank))) %>%
  group_by(set, cluster) %>%
  mutate(p_pop=n/sum(n)) %>%
  mutate(pp =.value * p_pop) %>%
  summarise(prev=sum(pp)) %>%
  group_by(cluster) %>%
  median_qi(prev)
c
```
And plot this - is there shrinkage?

So there is shrinkage, yes!

```{r}
# scale_clust <- scale %>% group_by(cluster) %>% summarise(n_pos=sum(hiv=="pos"), n=n()) %>% mutate(p_pos=n_pos/n) %>% filter(is.na(clust)==F)
# scale_clust$clust <- reorder(scale_clust$clust, scale_clust$p_pos)

ggplot() +
  #geom_point(aes(x=clust,y=p_pos,color="observed"), data=scale_clust) +
  geom_pointrange(aes(x=cluster, y=prev*100000, ymin=.lower*100000, ymax=.upper*100000, color="modelled"), data=c, size=0.2) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=-90))
```

## 8. Scaffold from the census

Same as the steps above, except this time, match to the census (wp) not the observed data.

HIV prevalence goes down, still a v small CI.

```{r}
newdata_wp <- wp_2019 %>% mutate(age=as.character(ageg))%>%
                mutate(age=factor(age,labels=age_levels_hack)) %>% 
                mutate(age=as.numeric(as.character(age))) %>% 
                mutate(cluster=as.character(cluster),sex = str_to_title(sex)) %>% 
                ungroup() %>% select(age,cluster,sex,N_wp) %>%
                rename(n=N_wp)

rm(b)
d <- tidybayes::add_fitted_draws(newdata=newdata_wp, model=m1) 
saveRDS(d, file = here::here("data/d.rds"))

d %>% ungroup() %>% 
  mutate(set=rep(1:4000, times=nrow(newdata_wp))) %>%
  group_by(set) %>%
  mutate(p_pop=n/sum(n)) %>%
  mutate(pp =.value * p_pop) %>%
  summarise(prev=sum(pp)) # So this is the overall prevalence, in 4000 different iterations of draws

d %>% ungroup() %>% 
  mutate(set=rep(1:4000, times=nrow(newdata_wp))) %>%
  group_by(set) %>%
  mutate(p_pop=n/sum(n)) %>%
  mutate(pp =.value * p_pop) %>%
  summarise(prev=sum(pp)) %>% # So this is the overall prevalence, in 4000 different iterations of draws
  median_qi(prev) # and median_qi it?

0.002536124*100000;	0.001708527*100000;	0.003619384*100000
```
And what about per cluster with census data...?

It goes up as expected

```{r}
e <- d %>% ungroup() %>% 
  mutate(set=rep(1:4000, times=nrow(newdata_blank))) %>%
  group_by(set, cluster) %>%
  mutate(p_pop=n/sum(n)) %>%
  mutate(pp =.value * p_pop) %>%
  summarise(prev=sum(pp)) %>%
  group_by(cluster) %>%
  median_qi(prev)

ggplot() +
  #geom_point(aes(x=clust,y=p_pos,color="observed"), data=scale_clust) +
  geom_pointrange(aes(x=cluster, y=prev*100000, ymin=.lower*100000, ymax=.upper*100000, color="modelled"), data=c, size=0.2) +
  geom_pointrange(aes(x=cluster, y=prev*100000, ymin=.lower*100000, ymax=.upper*100000, color="post-strat"), data=e, size=0.2) +
  #geom_hline(aes(color="observed", yintercept=median(scale_clust$p_pos)*100)) +
    #geom_hline(aes(color="modelled", yintercept=median(c$prev)*100)) +
    #geom_hline(aes(color="post-strat", yintercept=median(e$prev)*100)) +
 # scale_color_manual(values=c("#DD8D29","#46ACC8","#B40F20")) +
 # labs(y="HIV prevalence (%)",
  #     x="Cluster") +
  theme_bw() +
  theme(axis.text.x = element_blank())
```
# 9. Poststratify "by hand" to check

No modelling.  And ignoring cluster for now - just age and sex.

```{r}
f <- ind_xray_hts_subdata_tbcse %>% 
  mutate(age = cut(age, breaks=c(15,20,25,30,35,40,45,50,55,60,65,70,Inf), right=F,  labels=age_levels_hack)) %>%
  mutate(age=as.numeric(as.character(age))) %>%
  group_by(age,sex) %>% summarise(n_pos=sum(tbcase==1), n=n()) %>% 
  ungroup() %>%
  mutate(p_pos=n_pos/n,
         p_pop_tested=n/sum(n))

rm(d)
g <- newdata_wp %>% group_by(age,sex) %>% summarise(N_wp=sum(n)) %>% ungroup() %>% mutate(p_pop=N_wp/sum(N_wp)) %>% left_join(f)
saveRDS(g, file = here::here("data/g.rds"))

ggplot(g) +
  geom_path(aes(x=age, y=p_pop_tested, group=sex, color="Tested")) +
  geom_path(aes(x=age, y=p_pop, group=sex, color="WP population")) +
  facet_grid(.~sex) +
  theme_bw()
```

So men were undesampled -- and we've undersampled people with a relatively high risk of TB.  And blue is oversampled, and we've oversampled people with a relatively low risk of TB.

```{r}
g %>% select(age,sex,p_pos,p_pop,p_pop_tested) %>% 
  mutate(diff=p_pop_tested - p_pop) %>%
  gt() %>%
  fmt_number(columns = c(p_pos, p_pop_tested, p_pop, diff),rows = everything(),decimals = 3) %>%
  tab_style(
    style = list(
      cell_fill(color = "#8cd3ff")
      ),
    locations = cells_body(
      columns = c(diff,age,sex,p_pos),
      rows = diff > 0.025
    )) %>%
    tab_style(
    style = list(
      cell_fill(color = "#ffe5b4")
      ),
    locations = cells_body(
      columns = c(diff,age,sex,p_pos),
      rows = diff < -0.025
    )) 
```

```{r}
g %>% mutate(p1=p_pos * p_pop_tested) %>%
  mutate(p2=p_pos * p_pop) %>%
  summarise(sample=sum(p1),
            post_strat=sum(p2)) # goes down when post-stratify "by hand".

0.002149422*13492;0.002844623*13492

0.002149422*100000;0.002844623*100000

ind_xray_hts_subdata_tbcse %>% tabyl(tbcase)

```
## 10 Use the poststratified prevalances on the prev_tent data to calculate new prevalant TB cases
Import the scale dataset, use the column tent_cxr_total and the values in the tible named e
```{r}
dat_scale <- readRDS(here::here("data/dat_scale.rds"))
###make sure you check the calculation manually to verify what is going on
newdata_wp_cluster  <- newdata_wp %>% group_by(cluster) %>% summarise(n_cluster = sum(n))

dat_scale$n_cluster <- NULL
dat_scale$prev <- NULL

dat_scale <- dat_scale %>%
             left_join(newdata_wp_cluster, by = "cluster") %>%
             left_join(e, by = "cluster") %>%
             mutate(poststratPrevTB = n_cluster*prev, .after = n_prev_tbcases) %>%
             mutate(poststratPrevTB_round = round(poststratPrevTB), .after = poststratPrevTB)


dat_scale <- dat_scale %>%
                   mutate(poststratPrevTB_tent = tent_cxr_total*prev, .after = n_prev_tbcases) %>%
                   mutate(poststratPrevTB_round_tent = round(poststratPrevTB_tent), .after = poststratPrevTB)

dat_scale %>%
  select(cluster,contains("poststratPrevTB"),prev, worldpop=n_cluster) %>%
  gt::gt() %>%
  gt::fmt_missing(
    everything(),
    missing_text = ""
  ) %>%
  gt::gtsave(filename = "dat_scale_tent_poststratTB.rtf", path = here::here("figures"))

  
#View(dplyr::select(dat_scale, poststratPrevTB,poststratPrevTB_round,n_prev_tbcases, n_cluster, prev))

saveRDS(object = dat_scale, file = here::here("data/dat_scale.rds"))
```


<br>

## X. Reproducibility

This reproduction of the analysis was run by: 

```{r sysinfo, echo=FALSE, message=FALSE, comment=NA, warning=FALSE}

sysinfo <- Sys.info()

sysinfo <- data.frame(keyName=names(sysinfo), value=sysinfo, row.names=NULL)

sysinfo %>% kable()
```

Analysis was run at **`r Sys.time()`**, and using the following Session Info:

```{r sessioninfo, echo=FALSE, results='markdown', message=FALSE, comment=NA, warning=FALSE}
sessionInfo()
```
